Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2014-12-16 doc/GMP2015/s02_preliminaries.tex
Cut premilinary basics.

@@

A $d$-dimensional {\it geometric constraint system} in general consists of

- a underlying {\it constraint graph} $G=(V,E)$, or more generally, a hypergraph, whose  vertices represents geometric primitives in $d$-dimension and whose (hyper)edges specifies the combinatorics of constraints between pairs/tuples of the geometric primitives.

- an {\it algebraic system of  constraint functions} from $E$ to the domain of constraint values

In this paper we study qusecs, a class of 2D geometric constrain system that is independent (defined below).

A familiar example of geometric constraint system is {\it linkage (barjoint system)},
where the underlying constraint graph is a simple graph and
the constraints are Euclidean distances on the edges of the graph.

%--Give example as
%linkage(barjoint system) 
\todo{-- refer to others (citations
+ pinned line incidence and multi-body-pin (in application
section)}

%--underlying constraint graph or hypergraph (common treatment)
\todo{???The (hyper)graph of a is sometimes refered to???}

%--linearization: Jacobian of constraint system,
%rigidity matrix of indeterminates

The algebraic system can usually be linearized at {\it generic} points
by taking the Jacobian of the system,
giving  the $d$-dimensional {\it rigidity matrix } $R_d(G)$ of a the (hyper)graph $G = (V,E)$, 
whose rows correspond to edges, columns correspond to coordinate positions of vertices,  entries are \emph{indeterminates} representing the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$ of the geometric primitives corresponding to vertices $v \in V$.

%-- give example for the case of
%linkages

In the case of linkage, 
let $p_1(v), p_2(v), \ldots, p_d(v)$ represent the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$ of the joint corresponding to a vertex $v \in V$. 
The matrix $R_d(G)$ has one row for each edge $e \in E$ and $d$ columns for each vertex $v \in V$.  The row corresponding to $e = \{u,v\} \in E$ represents the bar connecting $p(u)$ and $p(v)$ and has $d$ non-zero indeterminate entries $\mathbf{p}(u)-\mathbf{p}(v)$ (resp.\ $\mathbf{p}(v)-\mathbf{p}(u)$) in the $d$ columns corresponding to $u$ (resp.  $v$) and zero in the other entries.

@@

A subset of edges $E'$, or a subgraph $(V', E')$, of a graph $G = (V, E)$ is said to be {\it independent} in $d$-dimensions, when the set of rows of $R_d(G)$ corresponding to $E'$ is generically independent, or independent for a generic instantiation of the indeterminate entries. 
This yields the $d$-dimensional generic {\it rigidity matroid} associated with a graph $G$.

The graph is  {\it infinitesimally rigid} if %the rigidity matrix has full rank, that is, 
if the number of generically independent rows or the rank of $R_d(G)$ is maximal.
In the case of linkage, the maximal rank is $d|V| - {d+1 \choose 2}$, where ${d+1 \choose 2}$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^d$ [?].
A vector in the kernel of the rigidity matrix is called an {\it infinitesimal motion vector}.

@@

The left nullspace of the rigidity matrix is called the {\it space of stress}, consisting of the {\it flex vectors}.

 A generalized version of Asimow Roth (implicit fn thm) states that
 generic rigidity, i.e.\ there exist at most finitely many solutions to the system at generic points,  %to the algebraic system 
  is equivalent to  our definition  of infinitesimal rigidity based on indeterminates.

@@

A graph $G = (V,E)$ is:
\begin{itemize}
	\item generically {\it isostatic (minimally rigid, well-constrained)} if $R_d(G)$ has full rank
	\item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
	\item {\it flexible} if $G$ is not rigid
    \item \textbf{under-constrained} if $G$ is independent and not rigid.
    \item \textbf{over-constrained} if it is not independent.
\end{itemize}
In this paper all results are about independent - underconstrained or
isostatic

\todo{--rigidity matroid\\
--Combinatorial characterization of rigidity matroid
Example: Laman's theorem for bar-joint (refer to others)
Lovasz Yemini
Tay white and whiteley
(citations)
+ pinned line incidence and multi-body-pin (in application
section)}

>>>>

In this paper, a {\em geometric constraint system} is a multivariate 
polynomial (usually bilinear or quadratic) system
$G(x,\delta)=0$, representing constraints with parameters $\delta$ 
between geometric primitives  in $\mathbb{R}^2$ represented collectively $x\in \mathbb{R}^n$.   
%
When the type of constraint (system) is fixed, 
the system is simply represented as $(G,\delta)$, where 
$G$ is the underlying constraint (hyper)graph $G = (V,E)$ 
with the vertices $V$ representing the geometric primitives in 
$\mathbb{R}^2$ and (hyper)edges
$E$ representing the constraints, each with an associated 
parameter $\delta$.

For example, a {\em bar-joint system or linkage} $(G,\delta)$, is a graph
$G=(V,E)$  
with fixed length bars as edges, i.e. $\delta: E \rightarrow \mathbb{R}$;
this represents the distance constraint system $\| x_u -x_v \|_2 =
\delta_{u,v}$ for  $(u,v) \in E$, where $x_u \in \mathbb{R}^2$ represents the coordinates of
$u\in V$.

In this paper, we will consider 3 types of constraint systems: bar-joint
as well as {\em pinned line-incidence}, and {\em body-hyperpin}. 
In all of these cases, a cartesian \emph{realization} or 
\emph{solution} $G(p)$ of $(G,\delta)$ is 
an assignment of coordinates or Euclidean transformations (poses), 
$p: V \rightarrow
\mathbb{R}^2$ or $\mathbb{R}^3$ to the vertices of $G$ satisfying the 
constraints with parameters $\delta$, 
modulo orientation preserving isometries 
(Euclidean rigid body motions except for pinned line-incidence).

Although the realization space itself depends on the constraint parameters 
$\delta$, many  relevant {\em generic} properties of 
the constraint system $G(x,\delta)$
are properties of the constraint (hyper)graph $G$ 
and do not depend on $\delta$ (or they hold for all but a measure 
zero set of $\delta$
values). Many of these are properties of the jacobian 
$\Delta_x G(x,\delta)$, often called the appropriate {\em rigidity matrix of
$G$} 
(a matrix of indeterminates).
For example, the {\em bar-joint rigidity matrix 
of the graph $G = (V,E)$}
is a matrix of indeterminates representing the 
jacobian of the distance map 
$\| x_u -x_v \|_2$  for $(u,v) \in E$.
The matrix has one row per edge $(u,v)$ in $E$,  
with 2 columns for $u$ (resp.
$v$) containing the 2 coordinate indeterminates 
for $x_u -x_v$ (resp. $x_v-x_u$), 
that is 4 non-zero entries per row.

Some of these generic constraint system or 
(hyper)graph properties are the following ({\bf Note:} we refer to these as
properties of the constraint system or as properties of the underlying
(hyper)graph interchangeably).
%
A constraint (hyper)graph $G$ is {\em independent} if its appropriate rigidity matrix of indeterminates  
has independent rows
(i.e, the determinant of some square submatrix is not
identically zero). 
It is {\em isostatic (minimally rigid, well-constrained)} 
if the number of generically independent rows or the rank of the appropriate rigidity matrix is maximal.
For example, the maximal rank of a bar-joint rigidity matrix is $2|V| - 3$, 
where $3$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^2$.
The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained. 
It is {\em flexible} if it is not rigid, {\em under-constrained} if it is independent and not rigid, or {\em over-constrained} if it is not independent.

In this paper, 
a {\em qusec} is any {\em independent} geometric 
constraint system of one of the 3
types mentioned above.

% % % % % % % % % % % % % % % % % % % % % % % % % % %

%A subset of edges $E'$, or a subgraph $(V', E')$, of a graph $G = (V, E)$ is said to be {\it independent} in $d$-dimensions, when the set of rows of $R_d(G)$ corresponding to $E'$ is generically independent, or independent for a generic instantiation of the indeterminate entries. 
%This yields the $d$-dimensional generic {\it rigidity matroid} associated with a graph $G$.

%The graph is  {\it infinitesimally rigid} if %the rigidity matrix has full rank, that is, 
%if the number of generically independent rows or the rank of $R_d(G)$ is maximal.
%In the case of linkage, the maximal rank is $d|V| - {d+1 \choose 2}$, where ${d+1 \choose 2}$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^d$ [?].
%A vector in the kernel of the rigidity matrix is called an {\it infinitesimal motion vector}.

%A graph $G = (V,E)$ is:
%\begin{itemize}
%	\item generically {\it isostatic (minimally rigid, well-constrained)} if $R_d(G)$ has full rank
%	\item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
%	\item {\it flexible} if $G$ is not rigid
%    \item \textbf{under-constrained} if $G$ is independent and not rigid.
%    \item \textbf{over-constrained} if it is not independent.
%\end{itemize}
%In this paper all results are about independent - underconstrained or
%isostatic

\todo{Where to put these}
A vector in the kernel of the rigidity matrix is called an {\it infinitesimal motion vector}.
The left nullspace of the rigidity matrix is called the {\it space of stress}. A vector in the stress space is called a {\it flex vector}.

% A generalized version of Asimow Roth (implicit fn thm) states that
% generic rigidity, i.e.\ there exist at most finitely many solutions to the system at generic points,  %to the algebraic system 
%  is equivalent to  our definition  of infinitesimal rigidity based on indeterminates.

%\todo{--rigidity matroid\\
%--Combinatorial characterization of rigidity matroid
%Example: Laman's theorem for bar-joint (refer to others)
%Lovasz Yemini
%Tay white and whiteley
%(citations)
%+ pinned line incidence and multi-body-pin (in application
%section)}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

@@

%A $d$-dimensional {\it geometric constraint system} in general consists of
%
%- a underlying {\it constraint graph} $G=(V,E)$, or more generally, a hypergraph, whose  vertices represents geometric primitives in $d$-dimension and whose (hyper)edges specifies the combinatorics of constraints between pairs/tuples of the geometric primitives.
%- an {\it algebraic system of  constraint functions} from $E$ to the domain of constraint value
%In this paper we study qusecs, a class of 2D geometric constrain system that is independent (defined below).

%--underlying constraint graph or hypergraph (common treatment)

%--linearization: Jacobian of constraint system,
%rigidity matrix of indeterminates
%
%The algebraic system can usually be linearized at {\it generic} points
%by taking the Jacobian of the system,
%giving  the $d$-dimensional {\it rigidity matrix } $R_d(G)$ of a the (hyper)graph $G = (V,E)$, 
%whose rows correspond to edges, columns correspond to coordinate positions of vertices,  entries are \emph{indeterminates} representing the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$ of the geometric primitives corresponding to vertices $v \in V$.

%-- give example for the case of
%linkages
%
%In the case of linkage, 
%let $p_1(v), p_2(v), \ldots, p_d(v)$ represent the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$ of the joint corresponding to a vertex $v \in V$. 
%The matrix $R_d(G)$ has one row for each edge $e \in E$ and $d$ columns for each vertex $v \in V$.  The row corresponding to $e = \{u,v\} \in E$ represents the bar connecting $p(u)$ and $p(v)$ and has $d$ non-zero indeterminate entries $\mathbf{p}(u)-\mathbf{p}(v)$ (resp.\ $\mathbf{p}(v)-\mathbf{p}(u)$) in the $d$ columns corresponding to $u$ (resp.  $v$) and zero in the other entries.

@@

@@

@@
